Einstein's gravity from a polynomial affine model

### 2016-05-11 AffGR-PRD.tex
Approved to be uploaded

%------------------
%--------- Definitions
%------------------
%% \input{Def-article.tex}
\usepackage{mathtools,amsmath,amssymb,amsfonts,dsfont,mathrsfs,amsthm}
\usepackage{graphicx}
\usepackage{centernot}
%% \usepackage{hyperref}
\usepackage{xcolor}
\usepackage{comment}
%% \hypersetup{linktocpage,colorlinks=true,urlcolor=blue!80!red,linkcolor=blue,citecolor=red}
%\usepackage{feynmf}
\usepackage{siunitx}
\usepackage{array}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
%% \usepackage{ulem}
%\usepackage{tikz}
\usepackage{braket}
\usepackage{xparse}

%-------------------------------Theorems
\newtheorem{Def}{Definition}[section]
\newtheorem{Thm}[Def]{Theorem}
\newtheorem{Lem}[Def]{Lemma}
\newtheorem{Pos}[Def]{Postulate}
\newtheorem{Exa}[Def]{Example}
\newtheorem{Cor}[Def]{Corrolary}
\newtheorem{Pro}[Def]{Proposition}

@@

\renewcommand{\(}{\left(}
\renewcommand{\)}{\right)}
\renewcommand{\[}{\left[}
\renewcommand{\]}{\right]}

@@

%% --------- Abstract
\begin{abstract}
  We show that the effective field equations for a recently formulated polynomial affine model of gravity, in the sector of a torsion-free connection, accept general Einstein manifolds---with or without cosmological constant---as solutions. Moreover, the effective field equations are partially those obtained from a gravitational Yang--Mills theory known as Stephenson--Kilmister--Yang theory. Additionally, we find a generalization of a minimally coupled massless scalar field in General Relativity within a ``minimally'' coupled scalar field in this affine model. Finally, we present a brief analysis of the propagators of the gravitational theory, and count the degrees of freedom. For completeness we include three appendices, including one in which we show that a Birkhoff-like theorem is valid for the analyzed sector.
\end{abstract}

\pacs{02.40.Ma,04.50.Kd,04.90.+e}
\keywords{Affine Gravity, Torsion, Generalized Gravity.}

@@

A hundred years ago, A.~Einstein presented by the first time the correct field equations of its relativistic theory of gravitation, known as General Relativity~\cite{einstein1915feldgleichungen}. It was not only consistent with the available observations, but it also predicted several effects that have been consistently observed and measured---see Ref.~\cite{Will:2014kxa} for a recent review---, being the last one the confirmation of the existence of the existence of gravitational waves~\cite{Abbott:2016blz}. Thus, General Relativity has become the most successful theory of gravitational interactions nowadays.

However, General Relativity seems to be an effective theory of a, yet to be discovered, more fundamental one. A first hint was pointed out by E.~Cartan, after remarking that the curvature is not built up from a metric field but from a connection, and that a part of it was left behind if one chooses to work with the Levi-Civita connection~\cite{Cartan1922,Cartan1923,Cartan1924,Cartan1925}. One could also see this by considering Palatini's approach in General Relativity, in which both metric and connection fields are independent, and noticing that generically the theory possesses terms which do not appear in General Relativity (when it is coupled to matter). In addition, we mention a few aspects related with the quantization: First, the fact that within the standard formalism of quantization, General Relativity is non-renormalizable~\cite{'tHooft:1973us,'tHooft:1974bx,Deser:1974cz,Deser:1974cy}, suggests the theory can be thought as an effective theory. Second, the Wheeler--DeWitt equations are not well-defined in general due to their nonpolynomial dependence on the Arnowitt--Deser--Misner (or ADM) variables~\cite{Arnowitt:1959ah,Arnowitt:1960es,WheelerGeo,DeWitt:1967yk,DeWitt:1967ub,DeWitt:1967uc}. One of the most successful quantization methods of gravitational interactions---developed from ideas by Gambini and Trias~\cite{Gambini:1980yz,Gambini:1986ew}, by Ashtekar~\cite{Ashtekar:1986yd,Ashtekar:1987gu}---, hints that General Relativity is not as complete as expected, since the Lagrangian formulation includes (at least) an extra term known as Holst term~\cite{Holst:1995pc}. Finally, it is well known that once matter couples to gravity (in particular fermionic matter), the assumption of vanishing torsion in standard General Relativity is no longer preserved as a field equation, and the best approach is to consider Cartan's generalization of General Relativity~\cite{Kibble:1961ba,Hehl:1976kj,Shapiro:2001rz,Hammond:2002rm}.
%%it is well known that in the standard formulation of General Relativity the skew-symmetric part of the connection (or torsion) is assumed to be zero, albeit it is known that such assumption can be relaxed once matter is added, in particular fermionic matter~\cite{Kibble:1961ba,Hehl:1976kj,Shapiro:2001rz,Hammond:2002rm}.

@@

Interestingly, gravitational theories with a generic linear connections have degrees of freedom which do not correspond to the expected massless graviton~\cite{Sezgin:1979zf}. Albeit these gravitational theories with an arbitrary connection possess ghosts, there are examples of well-behaved (stable) systems which have ghosts, such are the cases of the Pais--Uhlenbeck oscillator~\cite{Mannheim:2004qz,Bender:2007wu,Smilga:2008pr,Ilhan:2013xe}, or the higher derivative supersymmetric quantum mechanical system presented in Ref.~\cite{Robert:2008}. The last sentence, allows to argue that a model cannot be discarded  for the presence of ghosts model, without a rigorous analysis of its stability.

@@

Interestingly, by construction, the polynomial affine gravity proposed in Ref.~\cite{Skirzewski:2014eta} has no explicit terms leading to three-point graviton vertices, since all graviton \emph{self-interaction} is mediated by non-Riemannian parts of the connection. This characteristic might allow to bypass the general postulates supporting the no-go theorems stated in Refs.~\cite{McGady:2013sga,Camanho:2014apa}, where it was proved that generic three-point graviton interactions are highly constrained by causality and analyticity of the $S$-matrix, and the only \emph{acceptable} structure of the three-point graviton vertices is the one coming from General Relativity.
%%Additionally, it is well known that the three-point graviton vertices lead to non-renormalizability of a theory ---in a model independent way---~\cite{McGady:2013sga,Camanho:2014apa}. However, in the considered framework all graviton self-interaction is mediated by the torsion, which might allow to bypass the general postulates supporting the theorem.

The aim of this work is to show that in the vanishing torsion sector, the field equations of the polynomial affine gravity are a generalization of the Einstein's equations. % in vacuum. In addition, we show that ---in the same sector--- under the assumption of the existence of a unique non-degenerated, covariantly constant $\binom{0}{2}$-tensor, $g$ (which could be thought as a metric), the system of a scalar field coupled with the polynomial affine gravity accepts the minimally coupled Einstein--Klein--Gordon as a solution.
The paper is organized as follows. In Sec.~\ref{model}, we present briefly the polynomial affine gravity model, and summarize the results obtained in Ref.~\cite{Skirzewski:2014eta} concerning the Newtonian limit of the model. In Sec.~\ref{rlimit}, we restrict ourselves to the sector of the theory with vanishing torsion, and find the field equations; which surprisingly are a known generalization of the Einstein's field equation. Then, in Sec.~\ref{matter} we show that under certain considerations the coupling of the gravitational model with a scalar field represents a generalization of the minimally coupled system Einstein--Klein--Gordon. In Sec.~\ref{propag}, we present a brief analysis of the propagators of the model, which allows us to count the propagating degrees of freedom. In addition, we corroborate that result with a method equivalent to the one of Dirac. And in Sec.~\ref{conclusions} we summarize and discuss the possible implications of the results. In order to make the paper self-contained, we have included three appendices: In Appendix~\ref{DA}, we describe a dimensional analysis strategy that allowed us to build up the most general action consistent with diffeomorphisms invariance. In Appendix~\ref{Known}, we show a collection of known solutions to the effective field equations of the model. Finally, in Appendix~\ref{Schw}, we demonstrate---restricting ourselves to the vanishing torsion sector---the equivalent of the Birkhoff's theorem of the theory.

@@

First of all, we highlight that the model constructed below has as only fundamental field an affine connection, and no metric field is needed either for contracting nor lowering or raising indices. Moreover, in order to guarantee the correct transformation of the Lagrangian density, the geometrical objects used to write down the action will be the curvature and torsion of an affine connection, $\hat{\Gamma}^\mu{}_{\rho\sigma}$, which accepts a decomposition on irreducible components as

@@

  \hat{\Gamma}^\mu{}_{\rho\sigma} = {\Gamma}^\mu{}_{\rho\sigma} + T^\mu{}_{\rho\sigma} = {\Gamma}^\mu{}_{\rho\sigma} + \epsilon_{\rho\sigma\lambda\kappa}T^{\mu,\lambda\kappa}+A_{[\rho}\delta^\mu_{\nu]},

@@

where ${\Gamma}^\mu{}_{\rho\sigma}$ is symmetric in the lower indices, $A_\mu$ is a vector field corresponding to the trace of torsion, and  $T^{\mu,\lambda\kappa}$ is a Curtright field~\cite{Curtright:1980yk}, satisfying \mbox{$T^{\kappa,\mu\nu } = - T^{\kappa,\nu\mu }$} and $\epsilon_{\lambda\kappa\mu\nu}T^{\kappa,\mu\nu }=0$. Since no metric is present, the epsilon symbols are not related by raising (lowering) their indices, but instead we demand that \mbox{$\epsilon^{\delta\eta\lambda\kappa}\epsilon_{\mu\nu\rho\sigma}=4!\delta^{\delta}{}_{[\mu}\delta^\eta{}_{\nu}\delta^{\lambda}{}_{\rho} \delta^\kappa{}_{\sigma]}$.}

@@

In Eq.~\eqref{4dfull}, we have dropped all terms which can be related through partial integration, and those which are topological invariant (e.g. the Euler density). A straightforward dimensional analysis shows the impossibility of other contributions into the action, see Appendix~\ref{DA} (in particular Sec.~\ref{sec:ld}). Interestingly, the above action turns out to be power-counting renormalizable, which does not guarantee renormalizability but is a nice feature. The structure of the model yields no three-point graviton vertices, which might allow to overcome  the \emph{no-go} theorems found in Refs.~\cite{McGady:2013sga,Camanho:2014apa}. It is worth mentioning that all coupling constants are dimensionless, which might be a hint of conformal invariance of the model~\cite{Buchholz:1976hz}.

@@

First, we considered a static, homogeneous, and isotropic expansion of fields,
\begin{align}
  A_\mu &= \delta_\mu^0 A + a_\mu, \notag \\
  T^{\mu,\nu\rho} &= \delta^{\mu}_m\delta^{\nu\rho}_{m0}T + t^{\mu,\nu\rho},
  \label{GammaExp} \\
  \Gamma^\lambda{}_{\mu\nu} &= E \delta^\lambda_0 \delta^m_\mu \delta^m_\nu + F \delta^\lambda_m \delta^m_{(\mu}\delta^0_{\nu)} + G\delta^\lambda_0 \delta^0_{\mu}\delta^0_{\nu} + \gamma^\lambda{}_{\mu\nu}, \notag
\end{align}
with $\delta^{\mu\nu}_{\lambda\kappa}=\delta^{\mu}_{\lambda}\delta^{\nu}_{\kappa}-\delta^{\mu}_{\kappa}\delta^{\nu}_{\lambda}$. Next, by analysing the first order perturbations of the actions, we found a choice of the coupling constants which allow nontrivial solutions of these field equations, together with the non-relativistic limit of the geodesic equation. Then, we substitute the field components by their scalar perturbation decomposition,
\begin{equation}
  a_\mu \to \delta_\mu^0 a+\delta_\mu^m \partial_{m}a_{\mathtt{L}},
\end{equation}
\begin{equation}
  \begin{split}
    t^{\mu,\nu\rho} &\to \delta^{\mu}_m\delta^{\nu\rho}_{n0} \Big(t \delta^{m n} + \partial^m \partial^n t_{\mathtt{L}} \Big)
    +\delta^{\mu}_0 \delta^{\nu\rho}_{m0} \partial^m c_{\mathtt{L}}
    \\
    & \quad + \Big(\delta^{\mu}_0\delta^{\nu\rho}_{mn}-\delta^{\mu}_m\delta^{\nu\rho}_{n0}\Big)\epsilon^{m n p} \partial_{p} b
    \\
    & \quad +\delta^{\mu}_m \delta^{\nu}_{n} \delta^{\rho}_{p} \Big(\epsilon^{n p q}\partial_q \partial^m d_1 +  (\delta^{m n} \partial^p - \delta^{m p} \partial^n)d_2\Big)
  \end{split}
\end{equation}
and
\begin{equation}
  \begin{split}
    \gamma^\lambda_{\mu\nu}
    &\to
    \delta^\lambda_0\delta^0_\mu\delta^0_\nu u 
    + \delta^\lambda_m \delta^0_\mu\delta^0_\nu \partial^m v_{\mathtt{L}}
    + 2\delta^\lambda_0 \delta^0_{(\mu}\delta^m_{\nu)} \partial_m w_{\mathtt{L}}
    \\
    & \quad + \delta^\lambda_0 \delta^m_\mu\delta^n_\nu \Big(x \delta_{mn} + \partial_m \partial_n x_{\mathtt{L}}\Big)
    \\
    & \quad + 2\delta^\lambda_m \delta^0_{(\mu}\delta^n_{\nu)} \Big(y_1 \delta^m{}_n + \epsilon^{m p}{}_{n} \partial_p y_2 + \partial^m \partial_n y_{\mathtt{L}}\Big)
    \\
    & \quad + \delta^\lambda_m \delta^n_{\mu}\delta^p_{\nu} \Big(\delta_{n p} \partial^m z_1 + (\delta^m{}_n \partial_p+\delta^m{}_p \partial_n) z_2
    \\
    & \qquad +  (\epsilon^{m q}{}_n \partial_p+\epsilon^{m q}{}_p \partial_n) \partial_q z_3 + \partial^m \partial_n \partial_p z_{\mathtt{L}}\Big),
  \end{split}
\end{equation}
where the ``\texttt{L}'' sub-index identifies the longitudinal degrees of freedom.

@@

\section{\label{rlimit}Relativistic limit}
%%%%%%%%% COMPARISON GR %%%%%%%%%
Despite we obtained the correct Newtonian gravitational potential in the non-relativistic limit, see Eq.~\eqref{NewtonPot}, a \emph{post}-Newtonian approximation is necessary to explain gravitational phenomena like Mercury's perihelion and the deviation of light by a gravitational source. Therefore, we should go beyond the Newtonian limit in order to compare our model with General Relativity.

@@

%% Equations~\eqref{SimpleEOM} and~\eqref{harm-curv} are geometrically equivalent to the 
Notice that Eqs.~\eqref{SimpleEOM} and~\eqref{harm-curv} accept a geometrical interpretation equivalent to that of the field equations of a pure Yang--Mills theory, which in the language of differential forms are
\begin{align}
  \cdf \Ff*[2] &= 0, & \cdf \st \Ff*[2] &= 0,
\end{align}
where $\Ff*[2] = \cdf \Af*[1]$ is the field strength 2-form (the curvature 2-form of the connection in the principal bundle, see for example Ref.~\cite{bourguignon1982yang,Nakahara}), and the operator $\st$ denotes the Hodge star. Now, these Yang--Mills field equations are obtained from the variation of the action functional
\begin{equation}
  S_{\textsc{ym}} = \int \Tr \Big( \Ff*[2] \st \Ff*[2] \Big),
\end{equation}
and the Jacobi identity for the covariant derivative.

@@

A step toward the analysis of renormalizability of the model should be the computation of the free propagators of the model around a given background. The choice of background will necessarily in order to simplify the computations.  Our first task is to expand up to second order in the fields around a maximally symmetric Riemannian background. Maximally symmetric spaces possess remarkably simple curvature tensors~\cite{weinberg1972gravitation},

@@

      C_1 \big( - \Lambda {\nabla}_{\alpha}{{t}^{\beta \alpha \gamma}}\,  {g}_{\beta \gamma} - \Lambda {\gamma}^{\alpha}\,_{\beta \delta} {g}_{\alpha \gamma} {t}^{\beta \delta \gamma} + {\nabla}_{\alpha}{{\gamma}^{\alpha}\,_{\beta \delta}}\,  {\nabla}_{\gamma}{{t}^{\beta \delta \gamma}}\,  - {\nabla}_{\alpha}{{\gamma}^{\beta}\,_{\beta \delta}}\,  {\nabla}_{\gamma}{{t}^{\delta \alpha \gamma}}\, \big)

@@

In this paper we have shown that the \emph{polynomial affine gravity} presented in Ref.~\cite{Skirzewski:2014eta} is the most general (polynomial) model for an affine theory of gravity, see the Appendix~\ref{sec:ld}, consistent with invariance under diffeomorphisms. The action in Eq.~\eqref{4dfull}, is parametrized by a nineteen-dimensional moduli space---after the boundary and topological terms have been dropped---, which in some sense made us wonder about a possible \emph{duality} between this model and the one proposed in Ref.~\cite{Pagani:2015ema}, even though the last one possesses a metric field.
%% (boundary condition, topological terms, ...)

The field equations obtained from the action~\eqref{4dfull} are a set of very complex coupled partial differential equations. However, they are consistent with the limit of $T^{\lambda,\mu\nu}$ and $A_\mu$ going to zero. This consistency is what is usually called a \emph{consistent truncation} of the theory. In this paper we have restrict ourselves to such a truncation of our model, which constrained us to the torsion-less sector of the moduli space. Additionally, this truncation leaves only one nontrivial field equation, Eq.~\eqref{almostSimpleEOM}.
%%, which can be obtained from a simple ``effective'' theory whose action is Eq.~\eqref{eff-action}.
Moreover, by considering torsion-free equi-affine connections---those preserving a volume element---or a Levi-Civita connection---in the cases where the manifold is metric---, the remaining field equation simplifies even further to Eq.~\eqref{SimpleEOM}.

>>>>

%------------------
%--------- Definitions
%------------------
%% \input{Def-article.tex}
%% \usepackage{mathtools,amsmath,amssymb,amsfonts,dsfont,mathrsfs,amsthm}
%% \usepackage{graphicx}
%% \usepackage{centernot}
%% \usepackage{hyperref}
\usepackage{xcolor}
%% \usepackage{comment}
%% \hypersetup{linktocpage,colorlinks=true,urlcolor=blue!80!red,linkcolor=blue,citecolor=red}
%\usepackage{feynmf}
\usepackage{siunitx}
\usepackage{array}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
%% \usepackage{ulem}
%\usepackage{tikz}
\usepackage{braket}
\usepackage{xparse}

%-------------------------------Theorems
%% \newtheorem{Def}{Definition}[section]
%% \newtheorem{Thm}[Def]{Theorem}
%% \newtheorem{Lem}[Def]{Lemma}
%% \newtheorem{Pos}[Def]{Postulate}
%% \newtheorem{Exa}[Def]{Example}
%% \newtheorem{Cor}[Def]{Corrolary}
%% \newtheorem{Pro}[Def]{Proposition}

@@

%% \renewcommand{\(}{\left(}
%% \renewcommand{\)}{\right)}
%% \renewcommand{\[}{\left[}
%% \renewcommand{\]}{\right]}

@@

%% --------- Abstract
\begin{abstract}
  We show that the effective field equations for a recently formulated polynomial affine model of gravity, in the sector of a torsion-free connection, accept general Einstein manifolds---with or without cosmological constant---as solutions. Moreover, the effective field equations are partially those obtained from a gravitational Yang--Mills theory known as Stephenson--Kilmister--Yang theory. Additionally, we find a generalization of a minimally coupled massless scalar field in General Relativity within a ``minimally'' coupled scalar field in this affine model. Finally, we present a brief analysis of the propagators of the gravitational theory, and count the degrees of freedom. For completeness we prove that a Birkhoff-like theorem is valid for the analyzed sector.
\end{abstract}

\pacs{02.40.Hw,04.50.Kd,04.90.+e}
\keywords{Polynomial Affine Gravity, Torsion, Generalized Gravity.}

@@

A hundred years ago, A.~Einstein presented by the first time the correct field equations of its relativistic theory of gravitation, known as General Relativity~\cite{einstein1915feldgleichungen}. It was not only consistent with the available observations, but it also predicted several effects that have been consistently observed and measured---see Ref.~\cite{Will:2014kxa} for a recent review---, being the last one the confirmation of the existence of gravitational waves~\cite{Abbott:2016blz}. Thus, General Relativity has become the most successful theory of gravitational interactions nowadays.

However, General Relativity seems to be an effective theory of a, yet to be discovered, more fundamental description of gravitational effects. A first hint was pointed out by E.~Cartan, after remarking that the curvature is not built up from a metric field but from a connection, and that a part of it was left behind if one chooses to work with the Levi-Civita connection~\cite{Cartan1922,Cartan1923,Cartan1924,Cartan1925}. One could also see this by considering Palatini's approach in General Relativity, in which both metric and connection fields are independent, and noticing that generically the theory possesses terms which do not appear in General Relativity (when it is coupled to matter). In addition, we mention a few aspects related with the quantization: First, the fact that within the standard formalism of quantization, General Relativity is non-renormalizable~\cite{'tHooft:1973us,'tHooft:1974bx,Deser:1974cz,Deser:1974cy}, suggests the theory can be thought as an effective theory. Second, the Wheeler--DeWitt equations are not well-defined in general due to their nonpolynomial dependence on the Arnowitt--Deser--Misner (or ADM) variables~\cite{Arnowitt:1959ah,Arnowitt:1960es,WheelerGeo,DeWitt:1967yk,DeWitt:1967ub,DeWitt:1967uc}. One of the most successful quantization methods of gravitational interactions---developed from ideas by Gambini and Trias~\cite{Gambini:1980yz,Gambini:1986ew}, by Ashtekar~\cite{Ashtekar:1986yd,Ashtekar:1987gu}---, hints that a first order formulation of General Relativity is not as complete as expected, since the Lagrangian formulation should include an extra term known as Holst term~\cite{Holst:1995pc}. Finally, it is well known that once matter couples to gravity (in particular fermionic matter), the assumption of vanishing torsion in standard General Relativity is no longer preserved as a field equation, and the best approach is to consider Cartan's generalization of General Relativity~\cite{Kibble:1961ba,Hehl:1976kj,Shapiro:2001rz,Hammond:2002rm}.
%%it is well known that in the standard formulation of General Relativity the skew-symmetric part of the connection (or torsion) is assumed to be zero, albeit it is known that such assumption can be relaxed once matter is added, in particular fermionic matter~\cite{Kibble:1961ba,Hehl:1976kj,Shapiro:2001rz,Hammond:2002rm}.

@@

Interestingly, gravitational theories with a generic linear connections have degrees of freedom which do not correspond to the expected massless graviton~\cite{Sezgin:1979zf}. Albeit these gravitational theories with an arbitrary connection possess ghosts, there are examples of well-behaved (stable) systems which have ghosts, such are the cases of the Pais--Uhlenbeck oscillator~\cite{Mannheim:2004qz,Bender:2007wu,Smilga:2008pr,Ilhan:2013xe}, or the higher derivative supersymmetric quantum mechanical system presented in Ref.~\cite{Robert:2008}. The last sentence, allows to argue that a model cannot be disregarded for the presence of ghosts model, without a rigorous analysis of its stability.

@@

Interestingly, by construction, the polynomial affine gravity proposed in Ref.~\cite{Skirzewski:2014eta} has no explicit terms leading to three-point graviton vertices, since all graviton \emph{self-interaction} is mediated by non-Riemannian parts of the connection. This feature might allow to bypass the general postulates supporting the no-go theorems stated in Refs.~\cite{McGady:2013sga,Camanho:2014apa}, where it was proven that generic three-point graviton interactions are highly constrained by causality and analyticity of the $S$-matrix, and the only \emph{acceptable} structure of the three-point graviton vertices is the one coming from General Relativity.
%%Additionally, it is well known that the three-point graviton vertices lead to non-renormalizability of a theory ---in a model independent way---~\cite{McGady:2013sga,Camanho:2014apa}. However, in the considered framework all graviton self-interaction is mediated by the torsion, which might allow to bypass the general postulates supporting the theorem.

The aim of this work is to show that in the vanishing torsion sector, the field equations of the polynomial affine gravity are a generalization of the Einstein's equations. % in vacuum. In addition, we show that ---in the same sector--- under the assumption of the existence of a unique non-degenerated, covariantly constant $\binom{0}{2}$-tensor, $g$ (which could be thought as a metric), the system of a scalar field coupled with the polynomial affine gravity accepts the minimally coupled Einstein--Klein--Gordon as a solution.
The paper is organized as follows. In Sec.~\ref{model}, we present briefly the polynomial affine gravity model, and summarize the results obtained in Ref.~\cite{Skirzewski:2014eta} concerning the Newtonian limit of the model. In Sec.~\ref{rlimit}, we restrict ourselves to the sector of the theory with vanishing torsion, and find the field equations; which surprisingly are a known generalization of the Einstein's field equation. Then, in Sec.~\ref{matter} we show that under certain considerations the coupling of the gravitational model with a scalar field represents a generalization of the minimally coupled system Einstein--Klein--Gordon. In Sec.~\ref{propag}, we present a brief analysis of the propagators of the model, which allows us to count the propagating degrees of freedom. In addition, we corroborate that result with a method equivalent to the one of Dirac~\cite{Diaz:2014yua}. Finally, in Sec.~\ref{conclusions} we summarize and discuss the possible implications of the results. In order to make the paper self-contained, we have included three appendices: In Appendix~\ref{DA}, we describe a dimensional analysis strategy that allowed us to build up the most general action consistent with diffeomorphisms invariance. In Appendix~\ref{Known}, we show a collection of known solutions to the effective field equations of the model. Finally, in Appendix~\ref{Schw}, we demonstrate---restricting ourselves to the vanishing torsion sector---the equivalent of the Birkhoff's theorem of the theory.

@@

First of all, we highlight that the model constructed below has as only fundamental field an affine connection, and no metric field is needed neither for contracting nor lowering or raising indices. Moreover, in order to guarantee the correct transformation of the Lagrangian density, the geometrical objects used to write down the action will be the curvature and torsion of an affine connection, $\hat{\Gamma}^\mu{}_{\rho\sigma}$, which accepts a decomposition on irreducible components as

@@

  \hat{\Gamma}^\mu{}_{\rho\sigma} = \hat{\Gamma}^\mu{}_{(\rho\sigma)} + \hat{\Gamma}^\mu{}_{[\rho\sigma]} = {\Gamma}^\mu{}_{\rho\sigma} + \epsilon_{\rho\sigma\lambda\kappa}T^{\mu,\lambda\kappa}+A_{[\rho}\delta^\mu_{\nu]},

@@

where ${\Gamma}^\mu{}_{\rho\sigma} = \hat{\Gamma}^\mu{}_{(\rho\sigma)}$ is symmetric in the lower indices, $A_\mu$ is a vector field corresponding to the trace of torsion, and  $T^{\mu,\lambda\kappa}$ is a Curtright field~\cite{Curtright:1980yk}, satisfying \mbox{$T^{\kappa,\mu\nu } = - T^{\kappa,\nu\mu }$} and $\epsilon_{\lambda\kappa\mu\nu}T^{\kappa,\mu\nu }=0$. Since no metric is present, the epsilon symbols are not related by raising (lowering) their indices, but instead we demand that \mbox{$\epsilon^{\delta\eta\lambda\kappa}\epsilon_{\mu\nu\rho\sigma}=4!\delta^{\delta}{}_{[\mu}\delta^\eta{}_{\nu}\delta^{\lambda}{}_{\rho} \delta^\kappa{}_{\sigma]}$.}

@@

In Eq.~\eqref{4dfull}, we have dropped all terms which can be related through partial integration, and those which are topological invariant (e.g. the Euler density). A straightforward dimensional analysis shows the impossibility of other contributions into the action, see Appendix~\ref{DA} (in particular Sec.~\ref{sec:ld}). Interestingly, the above action turns out to be power-counting renormalizable, which does not guarantee renormalizability, but is a nice feature. The structure of the model yields no three-point graviton vertices, which might allow to overcome  the \emph{no-go} theorems found in Refs.~\cite{McGady:2013sga,Camanho:2014apa}. It is worth mentioning that all coupling constants are dimensionless, which might be a hint of conformal invariance of the model~\cite{Buchholz:1976hz}, and further analysis of this invariance will be presented in Ref.~\cite{OCF-future3}.

@@

First, we considered a static, homogeneous, and isotropic expansion of fields,
\begin{align}
  A_\mu &= \delta_\mu^0 A + a_\mu, \notag \\
  T^{\mu,\nu\rho} &= \delta^{\mu}_m\delta^{\nu\rho}_{m0}T + t^{\mu,\nu\rho},
  \label{GammaExp} \\
  \Gamma^\lambda{}_{\mu\nu} &= E \delta^\lambda_0 \delta^m_\mu \delta^m_\nu + F \delta^\lambda_m \delta^m_{(\mu}\delta^0_{\nu)} + G\delta^\lambda_0 \delta^0_{\mu}\delta^0_{\nu} + \gamma^\lambda{}_{\mu\nu}, \notag
\end{align}
with $\delta^{\mu\nu}_{\lambda\kappa}=\delta^{\mu}_{\lambda}\delta^{\nu}_{\kappa}-\delta^{\mu}_{\kappa}\delta^{\nu}_{\lambda}$. Next, by analysing the first order perturbations of the actions, we found a choice of the coupling constants which allows nontrivial solutions of these field equations, together with the non-relativistic limit of the geodesic equation. Then, we substitute the field components by their scalar perturbation decomposition,
\begin{equation}
  a_\mu \to \delta_\mu^0 a+\delta_\mu^m \partial_{m}a_{\mathtt{L}},
\end{equation}
\begin{equation}
  \begin{split}
    t^{\mu,\nu\rho} &\to \delta^{\mu}_m\delta^{\nu\rho}_{n0} \Big(t \delta^{m n} + \partial^m \partial^n t_{\mathtt{L}} \Big)
    +\delta^{\mu}_0 \delta^{\nu\rho}_{m0} \partial^m c_{\mathtt{L}}
    \\
    & \quad + \Big(\delta^{\mu}_0\delta^{\nu\rho}_{mn}-\delta^{\mu}_m\delta^{\nu\rho}_{n0}\Big)\epsilon^{m n p} \partial_{p} b
    \\
    & \quad +\delta^{\mu}_m \delta^{\nu}_{n} \delta^{\rho}_{p} \Big(\epsilon^{n p q}\partial_q \partial^m d_1 +  (\delta^{m n} \partial^p - \delta^{m p} \partial^n)d_2\Big)
  \end{split}
\end{equation}
and
\begin{equation}
  \begin{split}
    \gamma^\lambda_{\mu\nu}
    &\to
    \delta^\lambda_0\delta^0_\mu\delta^0_\nu u 
    + \delta^\lambda_m \delta^0_\mu\delta^0_\nu \partial^m v_{\mathtt{L}}
    + 2\delta^\lambda_0 \delta^0_{(\mu}\delta^m_{\nu)} \partial_m w_{\mathtt{L}}
    \\
    & \quad + \delta^\lambda_0 \delta^m_\mu\delta^n_\nu \Big(x \delta_{mn} + \partial_m \partial_n x_{\mathtt{L}}\Big)
    \\
    & \quad + 2\delta^\lambda_m \delta^0_{(\mu}\delta^n_{\nu)} \Big(y_1 \delta^m{}_n + \epsilon^{m p}{}_{n} \partial_p y_2 + \partial^m \partial_n y_{\mathtt{L}}\Big)
    \\
    & \quad + \delta^\lambda_m \delta^n_{\mu}\delta^p_{\nu} \Big(\delta_{n p} \partial^m z_1 + (\delta^m{}_n \partial_p+\delta^m{}_p \partial_n) z_2
    \\
    & \qquad +  (\epsilon^{m q}{}_n \partial_p+\epsilon^{m q}{}_p \partial_n) \partial_q z_3 + \partial^m \partial_n \partial_p z_{\mathtt{L}}\Big),
  \end{split}
\end{equation}
where the ``\texttt{L}'' sub-index identifies the longitudinal degrees of freedom.

@@

\section{\label{rlimit}Relativistic limit}
%%%%%%%%% COMPARISON GR %%%%%%%%%
Despite we obtained the correct Newtonian gravitational potential in the non-relativistic limit, see Eq.~\eqref{NewtonPot}, a \emph{post}-Newtonian approximation is necessary to explain gravitational phenomena like Mercury's perihelion and the bending of light by a gravitational source. Therefore, we should go beyond the Newtonian limit in order to compare our model with General Relativity.

@@

%% Equations~\eqref{SimpleEOM} and~\eqref{harm-curv} are geometrically equivalent to the 
Notice that Eqs.~\eqref{SimpleEOM} and~\eqref{harm-curv} accept a geometrical interpretation equivalent to that of the field equations of a pure Yang--Mills theory, which in the language of differential forms are
\begin{align}
  \cdf \Ff* &= 0, & \cdf \st \Ff* &= 0,
\end{align}
where $\Ff* = \cdf \Af*$ is the field strength 2-form (the curvature 2-form of the connection in the principal bundle, see for example Ref.~\cite{bourguignon1982yang,Nakahara}), and the operator $\st$ denotes the Hodge star. Now, these Yang--Mills field equations are obtained from the variation of the action functional
\begin{equation}
  S_{\textsc{ym}} = \int \Tr \Big( \Ff* \st \Ff* \Big),
\end{equation}
and the Jacobi identity for the covariant derivative.

@@

A step toward the analysis of renormalizability of the model should be the computation of the free propagators of the model around a given background. The choice of background is necessary in order to simplify the computations.  Our first task is to expand up to second order in the fields around a maximally symmetric Riemannian background. Maximally symmetric spaces possess remarkably simple curvature tensors~\cite{weinberg1972gravitation},

@@

      C_1 \big( - \Lambda {\gamma}^{\alpha}\,_{\beta \delta} {g}_{\alpha \gamma} {t}^{\beta \delta \gamma} + {\nabla}_{\alpha}{{\gamma}^{\alpha}\,_{\beta \delta}}\,  {\nabla}_{\gamma}{{t}^{\beta \delta \gamma}}\,  - {\nabla}_{\alpha}{{\gamma}^{\beta}\,_{\beta \delta}}\,  {\nabla}_{\gamma}{{t}^{\delta \alpha \gamma}}\, \big)

@@

In this paper we have shown that the \emph{polynomial affine gravity} presented in Ref.~\cite{Skirzewski:2014eta} is the most general (polynomial) model for an affine theory of gravity, see the Appendix~\ref{sec:ld}, consistent with invariance under diffeomorphisms. The action in Eq.~\eqref{4dfull}, is parametrized by a nineteen-dimensional moduli space---after the boundary and topological terms have been dropped---, which in some sense made us wonder about a possible \emph{duality} between this model and the one proposed in Ref.~\cite{Pagani:2015ema}.
%% (boundary condition, topological terms, ...)

The field equations obtained from the action~\eqref{4dfull} are a set of very complex coupled partial differential equations. However, the system is consistent with the limit of $T^{\lambda,\mu\nu}$ and $A_\mu$ going to zero. This consistency is what is usually called a \emph{consistent truncation} of the theory. In this paper we have restricted ourselves to such a truncation of our model, which constrained us to the torsion-less sector of the moduli space. Additionally, this truncation leaves only one nontrivial field equation, Eq.~\eqref{almostSimpleEOM}.
%%, which can be obtained from a simple ``effective'' theory whose action is Eq.~\eqref{eff-action}.
Moreover, by considering torsion-free equi-affine connections---those preserving a volume element---or a Levi-Civita connection---in the cases where the manifold is metric---, the remaining field equation simplifies even further to Eq.~\eqref{SimpleEOM}.
